### 2017-07-03 chapters/introduction/physiology.tex
Update latest

Deoxgenated blood is transported from the body trough the veins and to
the right atrium (RA). From the RA, blood flows to the right ventricle
(RV), before it is sent through the the pulmonary ateries(PA) and to the
lungs, where the bloods gets oxygenated. From the lungs, the blood
travels trough the pulmonary veins to the left atrium (LA), and
finally to the left ventricle (LV), before is is sent out to the body
again trough the aorta (Ao). A schamtic view of the blood flow
through the body is shown in Figure ~\ref{}

@@

The heart itself is located approximately in the center of your chest
with the \emph{apex} of heart angled down towards the left of the
body, and the \emph{base} located just behind the breastbone.

@@

The myocardium is made up by individual cardiac muscle cells, or
cardiac myocytes which branch and join neighbouring cell, and form a
complicated fibrous network which is often referred to as a
\emph{functional syncytium}; the number of active muscle fibers cannot
be varied from beat to beat.

Each myocyte is composed of bundles of myofibriles, which again
contain \emph{myofilaments}. The myofilaments cosists of a repeated pattern
of lines and bands which can be seen as a collection of individual
contrictile units called \emph{sarcomeres}.

\subsection{Cardiac pump function}

@@

>>>>

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.7\textwidth]{chapters/introduction/figures/heart_anatomy.pdf}
\caption{The heart anatomy. (Adapted and modified from Wapcaplet in Sodipodi).}
\label{fig:heart_anatomy}
\end{figure}

Deoxgenated blood is transported from the body trough the veins and to
the right atrium (RA). From the RA, blood flows to the right ventricle
(RV), before it is sent through the the pulmonary ateries(PA) and to the
lungs, where the bloods gets oxygenated. From the lungs, the blood
travels trough the pulmonary veins to the left atrium (LA), and
finally to the left ventricle (LV), before is is sent out to the body
again trough the aorta (Ao). A schamtic view of the blood flow
through the body is shown in Figure ~\ref{fig:heart_anatomy}

@@

The heart itself is located approximately in the center of the chest
with the \emph{apex} of heart angled down towards the left of the
body, and the \emph{base} located just behind the breastbone.

@@

The myocardium is made up by individual cardiac muscle cells, or
cardiac myocytes which branch and join neighbouring cell, and form a
complicated fibrous network which is often referred to as a
\emph{functional syncytium} meaning that the number of active muscle
fibers cannot be varied from beat to beat.

Each myocyte is composed of bundles of myofibriles, which again
contain \emph{myofilaments}. The myofilaments cosists of a repeated pattern
of lines and bands which can be seen as a collection of individual
contractile units called \emph{sarcomeres}. A simple sketch of a
sarcomere is shown in Figure \ref{fig:sarcomere}.

\begin{figure}[htbp]
  \centering
    \includegraphics[width=1.0\textwidth]{chapters/introduction/figures/Sarcomere.png}
\caption{Illustration of the sarcomere structure.}
\label{fig:sarcomere}
\end{figure}

These sarcomeres are composed thick and thin filaments that slides
back and forth when the muscle contract or relaxes. For that reason we
also refer to this model as the sliding-filamen model. 
The thin filaments are made up by a protein called \emph{actin} while
the thick filaments are made up by a protein called \emph{myosin}.
In order for the muscle to contract, calcium enters the cell after a
triggered action potential. These calcium ions binds to a complex called
troponin C located at the thin filments, which then exposes the actin to the
myosion head. When myosin binds to actin we say that a
\emph{cross-bridge} is formed between the thin and thick
filments. Now several models exist for the crossbridge cycle that
constitutes the sliding filament theory, but main steps includes the
use of ATP in order to trigger the \emph{power stroke} which makes the

\subsection{Passive and Active properties of the myocardium}
The myocardium is composed by small blood vessels that supply the
myocardial cells with oxygen. When the myocardium contracts this
perfused blood is then squeezed out, resulting in an overall loss of
volume. Such materials are referred to as compressible. When the
volume are preserved we refer to the material as incompressible.

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.5\textwidth]{chapters/introduction/figures/Hill_muscle_model.pdf}
\caption{The classical three-element Hill muscle model with one
  contractile element and two non-linear springs, one arranged in
  series and one parallel. }
\label{fig:hill_muscle_model}
\end{figure}

\subsection{Ventricular pumping function}
\label{sec:ventricular_pumping_function}
The left ventricle is by far the most studied chamber of the heart,
mainly because this is the chamber responsible for ejecting blood into the
body and therefore a diseased left ventricle could be fatal. A heart
beat can be described in many ways. One way is by looking at the
electrical propagation of current that travels through the heart and
activates the myocardium, leading to contraction. Such a diagram in
known as the \emph{electrocardiogram} (ECG). Another way to descrive
the cardiac cycle is by means of the LV pressure and volume. Plotting
the LV volume on the $x-$axis and the LV pressure on the $y-$axis
provides an intuitive representaion of the cardiac cycle known as the
\emph{PV-loop}, see Figure \ref{fig:pv_loop}.

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.8\textwidth]{chapters/introduction/figures/cardiac_cycle.png}
\caption{Showing the relation between pressure and volume in the left
  ventricle during the cardiac cycle, which the different phases and
  valvular events marked. To the left we show the a reduced version of
the classical Wiggers diagram with pressure and volume plotted against
time on the $x-$axis. To the right we show the pressure volume loop
with volume plotted on the $x-$axis and pressure on the $y-$axis. }
\label{fig:pv_loop}
\end{figure}

@@

\subsection{Summar of Cardiac Physiology}
The multiscale and multiphysics phenomena governening the mechanics of
the heart is complex. How the molecular dynamics should be coupled to
the electrophysiology at the cellular level, how the electrophysiology
should be coupled to the mechanics at the organ level and what type of
feedback mechanisms that should be included is far from fully
understood. In this thesis we limit ourself to the study of what is
happending on the organ level.

### 2017-07-03 chapters/introduction/summary.tex
Add latest version of thesis
Update latest

\section{Cardiac Anatomy and Physiology}

The heart is the muscular organ responsible for circulating the blood
thoughout the body.

Deoxgenated blood transported from the body trough the veins to the
right atrium (RA)

\subsection{Myocardial structure and morphology}

\subsection{Myocardial Contractility}

>>>>

\section{Summary of papers}

\subsection{Paper 1}

\subsection{Paper 2}

\subsection{Paper 3}

\subsection{Paper 4}

\section{Closing remarks and future directions}

A couple remarks regarding the limitations of the methods used in this
thesis shoud be made. Here we start by listing a couple of statements
which should be taken into considerations.

\begin{itemize}
\item \emph{The quality of features/biomarkers you can extract from a data-driven
model cannot be any better than the data used as input to the
model}. Most of the results presented in this thesis are based on data
obtained from clinical measurements of real patients.
\item \emph{The spatial resolution of the parameters should be
    reflected in the spatial resolution of the observations.} When
  trying to fit data that are spatially resolved at some level,
  choosing parameters that are resolved at a finer level should be
  done with caution. The continuity in the underlying physics as well
  as regularization techniques could be used to ...

\end{itemize}

During the work of this thesis, several questions still remains open
and would require

\begin{itemize}
\item The active model for the myocardium is has .. Degree of
  tranverse activation. Matching of strain data..
\item Identifiability of parameters... Uniqueness of
  solutions.. Amount of regularization..
\item Appropriate boundary conditions.. In paper three we saw big
  differences in the choice of boundary conditions. Especially, the
  magnitude of the stress seems to 
\item Coupling of electrophysiologigy and mechanics in an
  ajoint-based. data assimilation framework. 
\end{itemize}
